MaBoSS for HPC environments: Implementations of the continuous time Boolean model simulator for large CPU clusters and GPU accelerators

### 2024-05-13 Article.tex
Capitalized Boolean

Here, we present adaptations of MaBoSS~\cite{stoll2012continuous, stoll2017maboss} --- a stochastic Boolean simulator that performs estimations of state probability trajectories based on Gillespie stochastic simulation algorithm~\cite{gillespie1976general} -- to modern HPC computing architectures, which provide significant speedups of the computation, thus allowing scrutinization and analysis of much larger boolean models. In particular, the problem of properly quantifying low abundant phenotypes~\cite{gillespie2013perspective} can now be tackled by making more realistic the large number of simulation needed to cover the space of possible trajectories.
The main contributions comprise two new implementations of MaBoSS:
\begin{itemize}
    \item MaBoSS.GPU, a GPU-accelerated implementation of MaBoSS, which is designed to exploit the computational power of massively parallel GPU hardware.
    \item MaBoSS.MPI, a parallel implementation of MaBoSS which can scale to multinode environments, such as large CPU clusters.
\end{itemize}

>>>>

Here, we present adaptations of MaBoSS~\cite{stoll2012continuous, stoll2017maboss} --- a stochastic Boolean simulator that performs estimations of state probability trajectories based on Gillespie stochastic simulation algorithm~\cite{gillespie1976general} -- to modern HPC computing architectures, which provide significant speedups of the computation, thus allowing scrutinization and analysis of much larger Boolean models. In particular, the problem of properly quantifying low abundant phenotypes~\cite{gillespie2013perspective} can now be tackled by making more realistic the large number of simulation needed to cover the space of possible trajectories.
The main contributions comprise two new implementations of MaBoSS:
\begin{itemize}
    \item MaBoSS.GPU, a GPU-accelerated implementation of MaBoSS, which is designed to exploit the computational power of massively parallel GPU hardware.
    \item MaBoSS.MPI, a parallel implementation of MaBoSS which can scale to multinode environments, such as large CPU clusters.
\end{itemize}
